working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-02-20 Manuscript_draft/Manuscript_draft.Rmd
More changes in the intro

**Paragraph 3** Expanding ideas with examples

Incompatibility system is another plant trait that appears to play an important role in foreign pollen effect. Species that are self incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. However, note the large variability in mating systems across populations @whitehead2018 [...](). The type of incompatibility, (i.e. whether sporophytic or gametophytic) is related to the location of pollen recognition; sporophytic incompatibility relates to signaling at the stigma surface while gametophytic occurs within the style @barrett1988.  This later acting pollen recognition mechanism is associated with greater negative effect than sporophytic recognistion @ashman2013. [if you introduce this,I eexpect you will use it in your analysis, remove]() [Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018)- IB this is a side problem for you, mention just once, and maybe in discusion. Now it has a lot of weigth.](). Moreover, other traits such as number of pollen grains per flower and number of ovules have been traditionally associated with the type of incompatibility system where species with higher pollen ovule ratios are predicted to be xenogamous and species with low pollen ovule ratios autogamous (REF) . Selfer species are known to have a reduction of herkogamy (REF) and less pollen production per ovule (REF) which can be interpretated as a reduction of pollen exported into the community [so, pollen and ovules are important by themselfs, or only because correlated? Unclear what your point is](). Other morphological traits, like stigma size can be determinant for the total pollen quatity that a stigma can receive and therefore related to do that pollen size would also play an important role. Example with pollen here.

[This last bit of the paragraph is still under development](Jose)

**Paragraph 4** Maybe connect with paragrph above?

>>>>

Plant traits are crucial to understand heterospecific pollen effect, but here it is important to notice that these traits can be seen from a donor perpective or a recipient one. In first place, when the donor is considered pollen size and pollen allelopathy could be the two main predictors of effect. How pollen size and allepothaty are involved in heterospecific pollen effect remains unclear yet. For example, Herrera xxxx observed that small pollen is more likely to clogg the stigma, here we argue that although this is possible, the quantities of heterospecific pollen in nature are generally low as mentioned above and bigger pollen is generally related with higher pollen tube growth rate, which add another layer of complexity to the understanding of the overall importance of this trait. Moreover, pollen allelopathy is still an unclear field for pollination ecologist. When the pollen recipeint is taking into account style length, stigmatic area, flower morphology, number of ovules and incompatibility system seems to be the main predictors of effect. In the case of style length species with shorter styles will have greater heterospecific pollen effect due to pollen competition is more likely to occur and therefore ovule interference. Interestingly, differences of heterospecific pollen load has been shown to do not be influenced by flowr morphology (restricted or unrestricted), and the main predictor of the heterospecific pollen load was stigma size (REF). Therefore, if bigger stigmas can receive more hp pollen, a greater quantity of hp can be translated in a possible greater negative effect. For species that are self-incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018).

### 2019-02-20 Manuscript_draft/Manuscript_draft.tex
More changes in the intro

\textbf{Paragraph 3} Expanding ideas with examples

Incompatibility system is another plant trait that appears to play an
important role in foreign pollen effect. Species that are self
incompatible have stronger barriers to heterospecific pollen than
self-compatible species Ashman and Arceo-Gómez (2013). However, note the
large variability in mating systems across populations Whitehead et al.
(2018) \href{}{\ldots{}}. The type of incompatibility, (i.e.~whether
sporophytic or gametophytic) is related to the location of pollen
recognition; sporophytic incompatibility relates to signaling at the
stigma surface while gametophytic occurs within the style Barrett
(1988). This later acting pollen recognition mechanism is associated
with greater negative effect than sporophytic recognistion Ashman and
Arceo-Gómez (2013). \href{}{if you introduce this,I eexpect you will use
it in your analysis, remove} (Nonetheless, an effect of foreign pollen
is a bit obscured by the variability within species, however species
that are strong selfers or strong outcrossers have less variablity in
mating systems and predictions of effect could be more realistic (see
figure 1 from Whitehead et al. 2018)- IB this is a side problem for you,
mention just once, and maybe in discusion. Now it has a lot of
weigth.)(). Moreover, other traits such as number of pollen grains per
flower and number of ovules have been traditionally associated with the
type of incompatibility system where species with higher pollen ovule
ratios are predicted to be xenogamous and species with low pollen ovule
ratios autogamous (REF) . Selfer species are known to have a reduction
of herkogamy (REF) and less pollen production per ovule (REF) which can
be interpretated as a reduction of pollen exported into the community
\href{}{so, pollen and ovules are important by themselfs, or only
because correlated? Unclear what your point is}. Other morphological
traits, like stigma size can be determinant for the total pollen quatity
that a stigma can receive and therefore related to do that pollen size
would also play an important role. Example with pollen here.

\href{Jose}{This last bit of the paragraph is still under development}

\textbf{Paragraph 4} Maybe connect with paragrph above?

>>>>

Plant traits are crucial to understand heterospecific pollen effect, but
here it is important to notice that these traits can be seen from a
donor perpective or a recipient one. In first place, when the donor is
considered pollen size and pollen allelopathy could be the two main
predictors of effect. How pollen size and allepothaty are involved in
heterospecific pollen effect remains unclear yet. For example, Herrera
xxxx observed that small pollen is more likely to clogg the stigma, here
we argue that although this is possible, the quantities of
heterospecific pollen in nature are generally low as mentioned above and
bigger pollen is generally related with higher pollen tube growth rate,
which add another layer of complexity to the understanding of the
overall importance of this trait. Moreover, pollen allelopathy is still
an unclear field for pollination ecologist. When the pollen recipeint is
taking into account style length, stigmatic area, flower morphology,
number of ovules and incompatibility system seems to be the main
predictors of effect. In the case of style length species with shorter
styles will have greater heterospecific pollen effect due to pollen
competition is more likely to occur and therefore ovule interference.
Interestingly, differences of heterospecific pollen load has been shown
to do not be influenced by flowr morphology (restricted or
unrestricted), and the main predictor of the heterospecific pollen load
was stigma size (REF). Therefore, if bigger stigmas can receive more hp
pollen, a greater quantity of hp can be translated in a possible greater
negative effect. For species that are self-incompatible have stronger
barriers to heterospecific pollen than self-compatible species Ashman
and Arceo-Gómez (2013). Nonetheless, an effect of foreign pollen is a
bit obscured by the variability within species, however species that are
strong selfers or strong outcrossers have less variablity in mating
systems and predictions of effect could be more realistic (see figure 1
from Whitehead et al. (2018)).
